---
output: reprex::reprex_document
knit: reprex::reprex_render
---
## Variable assignment

Let's demonstrate variable assignment in R

```{r}
a <- 1
b <- 2
c <- 3

x <- a + b + c

x
 print(x)

 cat(a, b, c)
```

## Arithmetic operators

& is element wise
&& is on the whole vector


## Assignment operators

```{r}
# This is the left assignment
a <- 10

# Right assignment (Warning on it)
# 1 -> b

cat(a, b)
```

## Miscellaneous operators

```{r}
# : sequence of numbers
# %in% if a value belongs to a vector
# %'% matrix * t(matrix)

a <- 1:10
b <- 2
c <- 30

b %in% a
c %in% a

m1 <- matrix(c(1, 2, 3, 5, 6, 7), nrow = 2)

m1 %*% t(m1)
```

## Function

```{r}
your_function <- function(param1, param2) {
    # Function body
    x <- (param1 * param2)
    return(x)
}

your_function(2, 10)
```

## Data types r

### Numeric data type

```{r}
a <- 12
b <- 3.14159

class(b)
typeof(a)

a <- 12L
typeof(b)

a <- 1 + 2i
typeof(a)
```

## String manipulation

Either use print(), cat(a, b) or paste(a, b) to print string elements

```{r}
a <- "Hello"
b <- "KGP"
c <- "Talkie from Laxmi Kant"

print(c)

paste(a, b, sep = "", collapse = " ")
```

## Format display

```{r}
x <- 3.141592
y <- format(x, digits = 2)

typeof(a)
typeof(b)

as.double(y)

format(x, scientific = TRUE)
```

## Operators for vector creation

```{r}
a <- "kgp"
x <- 1:10
x
seq(1, 10)

# ?seq()

seq(1, 10, .1)

c("apple", "kgp", "talkie", 1, 3)
```

## Using [] operators to access vector elements

```{r}
# Unlike python or other traditionnal languages, the index starts
# at 1 in R
a[1]

a[c(1, 2)]

# We put "kgp" to FALSE so that it's not printed
# and it's repetitive following the boolean order
a[c(TRUE, FALSE)]
```
## Vectors

### Vector manipulation

```{r}
a <- 1:10
b <- seq(10, 1, -1)
# Arithmetic perfroms element-wise
a + b
a - b
a * b
a / b
```

### Vector recycling

```{r}
c <- 1:2
a
c

a + c
a - c
a * c
a / c
```

### Sorting

```{r}
sort((a / c), decreasing = TRUE)
```

## List R

## Simple list creation
```{r}
c("kgp", "talkie", 1)
x <- list("kgp", "talkie", 1, 1.12, 11, TRUE)

x[5]
```

### Named list

```{r}
x <- list("kgp", "talkie", 1, TRUE)
names(x) <- c("string1", "string2", "numeric", "boolean")

x$string1
x$numeric
```

### List manipulation and merging

```{r}
x$string1 <- "KGP"
x$numeric <- 10

x$boolean <- a
x <- list(a, c, 1, TRUE)
names(x) <- c("string1", "string2", "numeric", "boolean")
x

# Merging
x <- list("one", "two", "three")
y <- list(1, 2, 3)

z <- c(x, y)
z
```

### Conversion list to vector

R programming only works with matrices and vectors

```{r}
x <- list(1, 2, 3)
y <- list(1, 4, 9)

# z <- x * y
# this will produce and error : Non numeric argument ot binary operator
# to do it

z <- unlist(x)*unlist(y)
z
z <- as.list(z)
# and we have again a list
z
```

## Matrix

```{r}
a <- c(1, 2, 3, 4, 5, 6)
x <- matrix(a)
# The result is a column matrix
x

# To have it by row
x <- matrix(a, byrow = FALSE)
x

# by default its size is the number of rows
x <- matrix(a, nrow = 2, byrow = FALSE)
x

x <- matrix(a, nrow = 2, byrow = TRUE)
x
```
### Matrix operations

```{r}
a <- matrix(c(1:10), nrow = 3)
b <- matrix(seq(1, 80, 7), nrow = 3)

a
b

a[1, 2] # Row 1 column 2

a[1, ] # Whole row 1

b[, 2] # Whole column 2
```

Let's do some basic operations

```{r}
a + 1

a + b
a - b
a * b

```

## Arrays r 

```{r}
a <- c(1, 2, 3)
b <- c(4, 5, 5)
z <- array(c(a, b), c(2, 3, 1))
z
```
Let's add more complexities to our Arrays

```{r}
a <- c(1:9) # 9
b <- c(1, 4, 9, 16, 25, 36, 49, 64, 81) # 9

z <- array(c(a, b))
z # This is a 1D array

# Lets convert it to multiple dimensions
z <- array(c(a, b), c(3, 3, 2))
z
```

### Name of columns

```{r}
cols <- c("col1", "col2", "col3")
rows <- c("row1", "row2", "row3")
mats <- c("matrix1", "matrix2")

z <- array(c(a, b), c(3, 3, 2), dimnames = list(rows, cols, mats))
z

z[1]

# Value from 3rd row, all columns and 2nd matrix
z[3, , 2]

# For all columns and all matrices
z[3, , ]
```

## Factors in R
Factors are the data objects which are used to categorize
the data and store it as levels. They can store both strings and integers

```{r}
names <- c("sun", "mon", "tue", "sun", "mon", "fri", "mon")
is.factor(names)
is.vector(names)

l <- factor(names)
is.factor(l)

l

factor(l, levels = c("sun", "mon", "tue", "fri"))
```

Another method to create factor with gl

```{r}
gl(3, 4, labels = c("sun", "mon", "tue", "fri"), ordered = TRUE)
```
